---
title: "Report PeerAssessment1"
author: "Guillermo ATT"
date: "11 de enero de 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Code for reading in the data and/or processing the data

```{r Reading in the data}
library(ggplot2)
unzip("activity.zip")
data<-read.csv("activity.csv",stringsAsFactors = FALSE)
data$date<-as.Date(data$date,format="%Y-%m-%d")
data$interval<-as.factor(data$interval)
## The structure of the data frame is shown below
str(data)
```


## 2. Histogram of the total number of steps each day
```{r}
## We first load the neccesary library
library(ggplot2)
stepseachday<-tapply(data$steps,data$date,sum)
qplot(stepseachday,binwidth=500)+labs(title="Histogram of Steps Taken per Day",  x = "Number of Steps per Day", y = "Days Count")
str(stepseachday)
typeof(stepseachday)

```

## 3. Mean and median number of steps taken
```{r}
mean(stepseachday,na.rm=TRUE)
median(stepseachday,na.rm=TRUE)
```

## 4. Timeserie plot of the average number of steps taken for each 5 min interval
```{r}
steps_per_interval<-tapply(data$steps,data$interval,mean,na.rm=TRUE)
x<-unlist(attr(steps_per_interval,"dimnames"))
plot(x,steps_per_interval,
     type="l",xlab="Interval",ylab="Steps",
     main="Daily activity pattern for average number of steps")
```

## 5. The 5 min interval that, on average, contains the max number of steps
```{r}
max<- which.max(steps_per_interval)
names(steps_per_interval[max])
```
**`r names(steps_per_interval[max])`** is the 5 min interval with the maximum number of steps on average that are **`r steps_per_interval[max]`**

## 6. Code to describe and show a strategy for imputing missing data
For imputing missing data we can do it by assigning the mean value of the same interval.
First, we will calculate the total number of missing values
```{r}
missing<-sum(is.na(data$steps))
percentage<-mean(is.na(data$steps))
```
The number of missing values is **`r missing`**. This represents a **`r percentage`** % of the total amount of data.

We now proceed to impute the missing values with the mean values for the same intervals
```{r}
fill.value <- function(steps, interval) {
    filled <- NA
    if (!is.na(steps))
        filled <- c(steps)
    else
        filled <- (steps_per_interval[interval])
    return(filled)
}
filled.data <- data
filled.data$steps <- mapply(fill.value, filled.data$steps, filled.data$interval)
str(filled.data)
```
We can check that there are no missing values
```{r}
sum(is.na(filled.data$steps))
```


## 7. Histogram of the total number of steps taken each day after missing values are imputed
```{r}
total.steps <- tapply(filled.data$steps, filled.data$date, FUN=sum)
qplot(total.steps, binwidth=1000, xlab="total number of steps taken each day")

```

## 8. Panel plot comparing the average number of steps taken per 5 minute interval across weekdays and weekends
```{r}
weekdayorweekend <- function(date) {
    day <- weekdays(date)
    if (day %in% c("lunes", "martes", "miércoles", "jueves", "viernes"))
        return("weekday")
    else if (day %in% c("sábado", "domingo"))
        return("weekend")
    else
        stop("invalid date")
}
filled.data$day <- sapply(filled.data$date, FUN=weekdayorweekend)
table(filled.data$day)
str(filled.data)
```


```{r}
averages <- aggregate(steps ~ interval + day, data=filled.data, mean)
ggplot(averages, aes(interval, steps,group=1)) + geom_line() + facet_grid(day ~ .)+ scale_x_discrete(breaks=c(0,400,800,1200,1600,2000,2400)) +  xlab("5-minute interval") + ylab("Number of steps")
```